\chapter{Existing Software}
    This project can be compared to other projects on more levels:
    \begin{itemize}
        \item{RPython, Ecstatic (Ruby, \cite{madsen2007ecstatic}):} comparison to approaches used 
            for type analysis in other dynamic languages, e.g. Python, Ruby. 
            Probably more interesting from the academic point of view.
        \item{Weverca \cite{hauzarhunting}, Pixy\cite{jovanovic2006pixy}:} comparison to other tools that analyse PHP, e.g. Weverca, Pixy, 
            but those tools focus on different aims, namely security vulnerabilities 
            through user input, XSS attacks, etc.
        \item{NetBeans, Zend Studio:} comparison to real world tools that analyze PHP for the same 
            type of bugs (unused variable, use of unassigned variable, etc.), 
            but those tools usually do not provide any documentation and 
            any comparison would have to be based either on benchmarking and 
            evaluation or on detailed study of their source code, which would be 
            time consuming.
        \item{HipHop:} its type analysis is meant only for compiler back end, 
            so it tries to find only one type for a variable, if that fails, 
            it does not continue with analysis of that variable. 
            (As far as I understood from the source codes, because there 
            are no articles or documentation).
        \item{Phantm \cite{kneuss2010using}:} very similar project, but 
            uses Scala. What PHP versions it supports? Intra-procedural? 
            Integration into a compiler/interpreter.
    \end{itemize}    
