
\label{phpunittable}The following table contains all the issues 
reported by Control Flow for Phalanger analysis 
of PHPUnit code base. The issues with the same 
cause are grouped.
%\begin{center}
    \begin{longtable}{| l | l | l | p{6cm} |}
    
    \hline 
    \textbf{File} &   \textbf{Line}    &   \textbf{Category}       &   \textbf{Note}  \\ \hline 
    \endfirsthead
    
    %\multicolumn{4}{c}{{\bfseries Continued from previous page}} \\
    %\phpunittableheader{}
    \endhead

    %\hline \multicolumn{4}{|r|}{{Continued on next page}} \\ \hline
    \endfoot

    %\hline
    \endlastfoot    
    
    
    \path{F\TestCase.php}    &   1722    &   Style          &   Foreach used in form \code{foreach(\$array as \$key=>\$val)} although the \code{\$key} variable was not used anywhere. \\ \hline
    \path{F\TestCase.php}    &   1726    &   Style          &   Unused variable \code{\$iRecursiveCheck}. \\ \hline    
    \path{F\Assert.php}      &   1861    &   Actual Error   &   Discussion in the \wthesis. \\ \hline
    \path{F\Assert.php}      &   1960    &   Documentation  &   The routine code allows one of the arguments to be an \code{array} and works with it as such, but the documentation states that it can only be \code{boolean}. \\ \hline        
    \path{F\Assert.php}      &   1896    &   Documentation  &   \code{assertSelectEquals} method restricts its parameter type to be \code{integer} and is invoked with \code{boolean}. 
                                                                        However, the parameter value gets only forwarded to \code{convertSelectToTag}, which accepts any type (mixed). \\ \hline    
    \path{U\XML.php}              &   544     &   Documentation  &   \multirow{3}{6cm}{Missing \code{false} in return value documentation.} \\ \cline{1-3}    
    \path{F\TestSuite.php}  &   737     &   Documentation  &   \\ \cline{1-3}        
    \path{U\Test.php}             &   294     &   Documentation  &   \\ \hline
    \path{U\GlobalState.php}      &   351     &   False Positive &   \multirow{4}{6cm}{Built-in documentation error.} \\ \cline{1-3}    
    \path{U\Test.php}             &   360     &   False Positive &   \\ \cline{1-3}
    \path{T\Command.php}        &   791     &   False Positive &   \\ \cline{1-3}
    \path{T\Command.php}        &   745     &   False Positive &   \\ \hline
    \path{F\TestCase.php}    &   1941    &   Documentation  &   Field \code{mockObjectGenerator} is documented to have type \code{array}, 
                                                                        but value of type \code{MockObject\_Generator} 
                                                                        is assigned to it. The documentation should be updated. \\ \hline
    \path{U\Test.php}             &   46      &   False Positive &   Function \code{trait\_exists} is conditionally declared if it does not exist.
                                                                        It follows the same signature of actual built-in function \code{trait\_exists}, 
                                                                        but it has empty body, therefore it does not use its arguments. \\ \hline
    \path{F\C\Count.php} &   100      &  False positive &   \\ \hline                                                                        
    \path{F\C\Count.php} &   115      &  Documentation &  The documentation of method \code{getCountOf} states that it returns \code{boolean}, but 
                                                                            it returns \code{integer}.\\ \hline                                                         
    \path{F\C\Count.php} &   90      &  Style &   \multirow{1}{6cm}{Not all paths in a routine return a value.} \\ \cline{1-3}
    \path{F\C\IsType.php} &   127      &  Style &   \\ \hline
    \path{F\C\EMessage.php} &   69      &  False positive &   \\ \hline
    \path{F\C\IsAnything.php} &   76      &  False positive &  Unused parameter of a method that implements an interface. \\ \hline
    \path{F\C\IsJson.php} &   88      &  Documentation &  Method \code{determineJsonError} has wrong type documentation for one of its parameters. \\ \hline    
    \path{F\C\Count.php} &   100      &  False positive &   \\ \hline    
    \path{T\Command.php} &   339      &  Style &   Debatable: function \code{ini\_set} expects string, however, anything given to it is implicitly converted to string. \\ \hline
    \path{U\Configuration.php} &   991      &  Documentation &   The documentation of method \code{getInteger} states that the return type is \code{boolean}. \\ \hline
    \path{U\D\Logger.php} &   198      &  Documentation &   Wrong documentation of field's type. \\ \hline
    \path{U\T\ResultPrinter.php} &   206      &  False positive &   \\ \hline
    \end{longtable}
%\end{center}

Note: paths are shortened according to this schema:
\begin{itemize*}
    \item \path{F} -- \path{Framework}
    \item \path{F\C} -- \path{Framework\Constraint}
    \item \path{T} -- \path{TextUI}
    \item \path{U} -- \path{Util}
    \item \path{U\D} -- \path{Util\DeprecatedFeature}
    \item \path{U\T} -- \path{Util\TextDox}
    \item \path{F\C\EMessage.php} -- \path{F\C\ExceptionMessage.php}
\end{itemize*}